---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    highlight: pygments
    keep_md: yes
    theme: united
---

## Loading and preprocessing the data

Specify global options

```{r setoptions, warning=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE, warning = FALSE)
```

Setup the working directory

```{r workdir, echo=TRUE}
setwd("D:/Work/Copy/R Workspaces/Programs/R Markdown/coursera_rr_a01")
```

Let's load the required libraries

```{r loaddata, echo=TRUE}
library(dplyr)
library(ggplot2)
library(readr)
```

Let's import the data from the zip file directly and display the top 10
observations.

```{r importfile}
df <- read.csv(unz("activity.zip", "activity.csv"), header = T,
               stringsAsFactors = FALSE)

head(df)
```

Check the metadata and display some summary statistics

```{r displaymeta}
str(df)
summary(df)
```

The date column needs to be converted into date format from the character
format.

```{r preprocess}
df$date2 <- as.Date(df$date)
str(df)
```

Drop the character column and rename date2 to date.

```{r renamecols}
df$date <- NULL
colnames(df)[3] <- "date"
head(df)
```

## What is mean total number of steps taken per day?

Compute total number of steps taken per day

```{r averagesteps}
total_steps_per_day <- group_by(.data = df, date) %>%
    summarise(total_steps = sum(steps))

total_steps_per_day
```

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
