---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    highlight: pygments
    keep_md: yes
    theme: united
---

## Loading and preprocessing the data

Specify global options

```{r setoptions, warning=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE, warning = FALSE)
options(scipen = 7)
```

Setup the working directory

```{r workdir, echo=TRUE}
setwd("D:/Work/Copy/R Workspaces/Programs/R Markdown/coursera_rr_a01")
```

Let's load the required libraries

```{r loaddata, echo=TRUE}
library(dplyr)
library(ggplot2)
library(readr)
```

Let's import the data from the zip file directly and display the top 6
observations.

```{r importfile}
df <- read.csv(unz("activity.zip", "activity.csv"), header = T,
               stringsAsFactors = FALSE)

head(df)
```

Check the metadata and display some summary statistics

```{r displaymeta}
str(df)
summary(df)
```

The date column needs to be converted into date format from the character
format.

```{r preprocess}
df$date2 <- as.Date(df$date)
str(df)
```

Drop the character column and rename date2 to date.

```{r renamecols}
df$date <- NULL
colnames(df)[3] <- "date"
head(df)
```

## What is mean total number of steps taken per day?

Compute total number of steps taken per day and save it in a new dataset.

```{r averagesteps}
total_steps_per_day <- group_by(.data = df, date) %>%
    summarise(total_steps = sum(steps, na.rm = TRUE))

total_steps_per_day
```

Generate a histogram of steps taken per day.

```{r histogram, fig.height=4, fig.width=4}
ggplot(total_steps_per_day, aes(total_steps)) +
    geom_histogram(binwidth = 3000, color = 'black',
                   fill = 'firebrick') +
    ggtitle('Steps Per Day') +
    theme(plot.title = element_text(size = 15, face = 'bold',
                                    vjust = 2)) +
    labs(x = "Total Steps", y = "Count")
```

Lets compute the mean and median number of steps taken per day.

```{r avgsteps}
avg_steps <- round(mean(total_steps_per_day$total_steps, na.rm = TRUE), 1)
avg_steps
med_steps <- round(median(total_steps_per_day$total_steps, na.rm = TRUE), 1)
med_steps
```

The average number of steps taken per day is **`r avg_steps`** and the median
number of steps taken per day is **`r med_steps`**.

## What is the average daily activity pattern?

Aggregate total steps taken at 5 min intervals averaged across days

```{r avgdailypattern}
avg_steps_per_interval <- group_by(.data = df, interval) %>%
    summarise(avg_steps = mean(steps, na.rm = TRUE))

# Look at top few rows
avg_steps_per_interval
```

Generate a time series plot of average number of steps taken over 5-minute
intervals.

```{r timeplot, fig.height = 4, fig.width = 6}
ggplot(avg_steps_per_interval, aes(x = interval, y = avg_steps)) +
    geom_line(colour = "firebrick") + 
    ggtitle("Avg Steps Over 5 Min Interval") +
    theme(plot.title = element_text(size = 15, face = 'bold',
                                    vjust = 2)) +
    labs(x = "Intervals", y = "Avg Steps")
```

Let's compute which 5-minute interval on average has the maximum number of 
steps

```{r computemax}
max_step_row <- top_n(x = avg_steps_per_interval, 1, wt = avg_steps)

max_step_row$interval
```

The 5-minute interval with the maximum number of steps on an average is 
`r max_step_row$interval`.

## Imputing missing values

Compute the number of rows with missing values.

```{r computemissing}
# Computing only for steps, since interal and date does not have missing
nrow_miss <- length(which(is.na(df$steps)))

nrow_miss
```

There are `r nrow_miss` rows missing value. We can try imputing the missing
with average of the corresponding 5-minute intervals across all other days.
We have already calculated the average in the table **avg_steps_per_interal**.
We can merge **avg_steps_per_interal** with **df** on **interval** and replace
missing values with **avg_steps**.

```{r imputemissing}
df2 <- df %>% left_join(avg_steps_per_interval)

# Check the output dataset
head(df2)

# Replace missing steps with avg_steps
df2$steps <- ifelse(is.na(df2$steps), df2$avg_steps, df2$steps)

# Check the output dataset
head(df2)

# Check if any missing value remains
which(is.na(df2))

# Drop avg_steps
df2$avg_steps <- NULL
```

Recomputing total steps per day, generating histograms and computing mean and 
median of total number of steps taken per day.

```{r totalsteps, fig.height=4, fig.width=4}
# Compute total steps
total_steps_per_day2 <- group_by(.data = df2, date) %>%
    summarise(total_steps = sum(steps, na.rm = TRUE))

# Display top 10 rows
total_steps_per_day2

# Generate a histogram of steps taken per day.
ggplot(total_steps_per_day2, aes(total_steps)) +
    geom_histogram(binwidth = 3000, color = 'black',
                   fill = 'firebrick') +
    ggtitle('Steps Per Day with Imputation') +
    theme(plot.title = element_text(size = 15, face = 'bold',
                                    vjust = 2)) +
    labs(x = "Total Steps", y = "Count")

# Lets recompute the mean and median number of steps taken per day.
avg_steps2 <- round(mean(total_steps_per_day2$total_steps, na.rm = TRUE), 1)
avg_steps2

med_steps2 <- round(median(total_steps_per_day2$total_steps, na.rm = TRUE), 1)
med_steps2
```

The earlier mean and median values of number of steps taken per day were 
**`r avg_steps`** and **`r med_steps`** respectively. Due to imputation, the 
mean and the median number of steps per day have increased to **`r avg_steps2`**
and **`r med_steps2`** respectively. The distribution of total number of steps
per day have become more symmetric around the mean. The distribution has become
more peaked as all the missing values are now appearing at the middle.

## Are there differences in activity patterns between weekdays and weekends?
